---
title: "hw03-gapminder"
output: github_document
---
## Including packages:
```{r}
library(gapminder)
library(tidyverse)
```
## Get the maximum and minimum of GDP per capita for all continents:
```{r}
gdpByCont <- gapminder %>% 
  group_by(continent) %>% 
  summarize(gdpPercap_min = min(gdpPercap), gdpPercap_max = max(gdpPercap))
gdpByCont
```
And then make a box plot to show the data above:
```{r}
gdpByCont %>% 
  # gather the max and min data in one "value"
  gather(variable,value,-continent) %>% 
  ggplot(aes(x=continent, y=value, fill = variable)) + 
  # make the max and min value adopt "dodge" positions in one bar plot
  geom_col(position = position_dodge())
  
```
## Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.
```{r}
wtMean <- gapminder %>% 
  group_by(continent,year) %>% 
  summarise(lifeExp=weighted.mean(lifeExp,pop))
wtMean
```
```{r}
ggplot(wtMean, aes(year,lifeExp)) +
  geom_smooth(method = 'loess',aes(colour=continent))
```

## How is life expectancy changing over time on different continents?
```{r}
gapminder %>% 
  group_by(continent,year) %>% 
  summarise(lifeExp=mean(lifeExp))
```

```{r}
ggplot(gapminder, aes(year,lifeExp)) +
  geom_point(aes(colour=continent),alpha=0.1) +
  geom_smooth(method = 'loess',aes(colour=continent))
```
## Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.
```{r}
popMean <- gapminder %>% 
  group_by(country,year,continent) %>% 
  filter(pop > 50000000) %>%  
  summarise(pop=mean(pop))
popMean
```
```{r}
ggplot(subset(popMean),
       aes(x = year, y = pop, group = country, color = country))+
       geom_line()
```
```{r}
# Mean <- gapminder %>% 
#   group_by(country,year,continent) %>% 
#   filter(gdpPercap > 20000) %>%  
#   summarise(gdpPercap=mean(gdpPercap))
# Mean
# ggplot(subset(Mean),
#        aes(x = year, y = gdpPercap, group = country, color = country))+
#        geom_line()
Mean <- gapminder %>% 
  group_by(country,year) %>% 
  filter(country == "Kuwait") %>%  
  summarise(gdpPercap=mean(gdpPercap))
Mean
ggplot(subset(Mean),
       aes(x = year, y = gdpPercap, group = country, color = country))+
       geom_line()
```

