---
title: "hw03-gapminder"
output: github_document
---
## Including packages:
```{r}
library(gapminder)
library(tidyverse)
```
## Get the maximum and minimum of GDP per capita for all continents:
```{r}
gdpByCont <- gapminder %>% 
  group_by(continent) %>% 
  summarize(gdpPercap_min = min(gdpPercap), gdpPercap_max = max(gdpPercap))
gdpByCont
```
And then make a box plot to show the data above:
```{r}
gdpByCont %>% 
  # gather the max and min data in one "value"
  gather(variable,value,-continent) %>% 
  ggplot(aes(x=continent, y=value, fill = variable)) + 
  # make the max and min value adopt "dodge" positions in one bar plot
  geom_col(position = position_dodge())
  
```

## Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. 
I computed the wieghted mean with the weighted.mean() function:
```{r}
wtMean <- gapminder %>% 
  group_by(continent,year) %>% 
  summarise(lifeExp=weighted.mean(lifeExp,pop))
wtMean
```
Then I tend to plot the results above in chronical order, and make the line smooth, coloured with continent:
```{r}
ggplot(wtMean, aes(year,lifeExp)) +
  geom_smooth(method = 'loess',aes(colour=continent))
```

## How is life expectancy changing over time on different continents?
First make a table to show the lifeExp values:
```{r}
gapminder %>% 
  group_by(continent,year) %>% 
  summarise(lifeExp=mean(lifeExp))
```
This time, I plot both the points and the smoothed line to show the lifeExp change in one graph:
```{r}
ggplot(gapminder, aes(year,lifeExp)) +
  geom_point(aes(colour=continent),alpha=0.1) +
  geom_smooth(method = 'loess',aes(colour=continent))
```
I've noticed that the lifeExp in Oceania tends to grow faster over time. In Enrope and America, the lifeExp grows smoothly. While in Asia and Africa, the growing speed become slower over time. 

## Find countries with interesting stories. Open-ended and, therefore, hard. Promising but unsuccessful attempts are encouraged.

Let's first focus on the country with population larger than 50 millions, with a table with mean population of the filtered countries:
```{r}
popMean <- gapminder %>% 
  group_by(country,year,continent) %>% 
  filter(pop > 50000000) %>%  
  summarise(pop=mean(pop))
popMean
```
NOw making a plot of the popolation growth in different countries:
```{r}
ggplot(subset(popMean),
       aes(x = year, y = pop, group = country, color = country))+
       geom_line()
```
The graph shows two specific countries with a high population growing speed: China and India. 

Then I'm trying to do something with the 'gdpPercap' variable. First make a table to show the values with gdpPercap larger than 20000:
```{r}
Mean <- gapminder %>%
  group_by(country,year,continent) %>%
  filter(gdpPercap > 20000) %>%
  summarise(gdpPercap=gdpPercap)
Mean
```
I then plot the data with geom_line:
```{r}
ggplot(subset(Mean),
       aes(x = year, y = gdpPercap, group = country, color = country))+
       geom_line()
```
From which I found the gdpPercap in Kuwait was very high at first, how ever it changed dramatically afterwards and went trough a rapidly drop between 1970s and 1980s. I did some searches and found from [Kuwait’s Souk al-Manakh Stock Bubble](http://www.thebubblebubble.com/souk-al-manakh/) that in this period it 
> experienced a major economic crisis after the Souk Al-Manakh stock market crash and decrease in oil price.

##But I want to do more:
I want to get table and figure side-by-side. I found the grid.arrage() function, which is in the 'gridExtra' library to make the layout in a grid view. For this situation, I just need to make the table and plot in 2 different columns.
(I reused the max-min 'gdpByCont' table from the first section.)
```{r} 
library("gridExtra")
# turn the table into tableGrob
t <-  tableGrob(gdpByCont)
# same operation as before to generate a plot
p <- gdpByCont %>% 
  gather(variable,value,-continent) %>% 
  ggplot(aes(x=continent, y=value, fill = variable)) + 
  geom_col(position = position_dodge())
# arrage them in 2-column grid view
grid.arrange(p, t, ncol=2)
```


